# Collection {#collect}

Data collection is the process of collecting data. 
Before construction of a data collection protocol, 
the reader should be familiar with future aspects of the data management 
pipeline and thus I suggest reading (or at least skimming) through the remainder
of the book in order to fully understand the suggests given here. 

## Methods

There are three main methods of data collection

  1. lab notebook or field sheets,
  1. direct digital entry, and
  1. automated digital file.
  
In lab notebooks or field sheets the data are written down on paper and 
will need to be entered into a computer and saved as a digital file later. 
In direct digital entry, a data recorded inputs the data directly into a 
computer and saves the results as a digital file, 
but does not have an original paper copy.
Finally, an automated digital file is created by a machine connected to 
a computer and the machine automatically creates a digital file.
  
For each of these methods,

  
### Lab notebook or field sheets

In field and laboratory settings, 
data are often recorded using pen on paper. 
These data are digitized and stored in a file on a computer.

#### Construction of lab notebook or field sheets

To make the lab notebook and field sheet amenable to digitization,



### Direct digital entry



### Automated digital file


## Data standards {#standards}

When writing data in a lab notebook or field sheet, 
directly into a digital file, or through an automated digital file, 
decisions must be made about the format to store the information.
The key here is not so much to make the correct decision, but, instead,
to be consistent in that decision.
So here are my suggestions for how to store data.

### Variable names

I suggest that variable names follow the 
[R convention for variable names](https://stat.ethz.ch/R-manual/R-devel/library/base/html/make.names.html)
which can be found here:
 
```{r variable_names, echo=TRUE, eval=FALSE}
?make.names
```

In particular, a valid name consists of letters, 
numbers, a period, and underscore characters.
The name must start either with a letter or 
a dot that is not followed by a number.
The name must not be a 
[reserved word](https://stat.ethz.ch/R-manual/R-devel/library/base/html/Reserved.html).

In addition, I suggest the use of lowercase letters except for the [key](#key)

**Rationale:** Lowercase letters are easier to type since you don't need the 
shift key.

### 

```{r data}
d <- as.data.frame(HairEyeColor)
names(d) <- tolower(names(d))
head(d)
```

### Include a key {#key}

The vast majority of data tables should have an associated 
[key](https://en.wikipedia.org/wiki/Unique_key). 
The key is an identifier to a record that links the record across different
data tables.
I identify this IDentifier using ID at the end of the variable name.

### Binary data

Be explicit about the two levels of binary data, 
rather than requiring a lookup table. 

For example, a variable called `political party` with levels `Democrat` and 
`Republican` should be stored exactly as written, e.g.


