# Variables



## Data standards {#standards}

When writing data in a lab notebook or field sheet, 
directly into a digital file, or through an automated digital file, 
decisions must be made about the format to store the information.
The key here is not so much to make the correct decision, but, instead,
to **be consistent** in that decision.
So here are my suggestions for how to store data.

### Variable names

I suggest that variable names follow the 
[R convention for variable names](https://stat.ethz.ch/R-manual/R-devel/library/base/html/make.names.html)
which can be found here:
 
```{r variable_names, echo=TRUE, eval=FALSE}
?make.names
```

In particular, a valid name consists of letters, 
numbers, periods ("."), and underscores ("_").
The name must start either with a letter or 
a dot that is not followed by a number.
The name must not be a 
[reserved word](https://stat.ethz.ch/R-manual/R-devel/library/base/html/Reserved.html).

In addition, I suggest the use of lowercase letters except for the [key](#key)

**Rationale:** Lowercase letters are easier to type since you don't need the 
shift key.

#### Example

In the following data set, 
the variable names are the column names and each is constructed from
lower case letters. 

```{r data}
d <- as.data.frame(HairEyeColor)
names(d) <- tolower(names(d))
names(d)[4] <- "count"
d
```



### Categorical data

Be explicit about the levels of categorical data, 
rather than requiring a lookup table. 

#### Example

Prefer 

```{r}
head(d)
```

to

```{r, echo=FALSE}
d2 <- d

d2$hair <- as.numeric(d2$hair)
d2$eye  <- as.numeric(d2$eye)
d2$sex  <- as.numeric(d2$sex)
```

```{r}
head(d2)
```

where the lookup tables are given below

```{r}
levels(d$hair)
levels(d$eye)
levels(d$sex)
```


### Dates should be in YYYY-MM-DD format

The important part here is that the date has the 4-digit year followed by the 
2-digit month followed by the 2-digit day, i.e. YYYYMMDD order.
It is really fine to use any separator (or none),
but if you use "/" or "-", then R can easily read the date.

So, the dates should look like

```{r, echo=FALSE}
as.Date(Sys.time())
```

**Rationale:** This is the 
[ISO 8601 standard](https://en.wikipedia.org/wiki/ISO_8601) and is immediately 
recognized by R.

If no day is available, then YYYY/MM will suffice and if no month is available
then YYYY will suffice.

If you want more detailed time, e.g. hours, minutes, and seconds, 
then record using YYYY/MM/DD HH:MM:SS using 24-hour time. 
The time zone could also be included to reduce confusion.
For example,

```{r, echo=FALSE}
Sys.time()
```




### Include a key {#key}

In order to record each piece of data only once, 
data tables will need to be constructed such that they can borrow information
in the future. 
In order for records in different tables to be able to reference each other
each data table should have a column that contains a
[key](https://en.wikipedia.org/wiki/Unique_key) for each record.
The key is an identifier to a record that links the record across different
data tables.
I identify this IDentifier using ID at the end of the variable name where
ID is intentionally written in capital letters to make it stand out from the
other variable names.


