# Scripts

Scripts will be utilized to perform a number of tasks once the raw data have 
been entered including 
[collating](#collating-scripts),
[joining](#joining-scripts),
[summarizing](#summarizing-scripts), and
[analyzing](#analyzing-scripts).
These scripts will be written in R.


## Collating scripts

As we are promoting a ``lots of little files'' approach, 
scripts will be needed to combine these files into a single `data.frame` for 
analysis in R.
Since we have made sure all our data files, within a data type directory), have
the same structure, it will be relatively easy to read all the files in and 
combine them. 
My generic code to perform this is available at this 
[gist](https://gist.github.com/jarad/8f3b79b33489828ab8244e82a4a0c5b3)
which currently contains this code

```{r read_dir, eval=FALSE}
library(dplyr)
library(tidyr)

my_read_csv = function(f, into) {
  readr::read_csv(f) %>%
    mutate(file=f) %>%
    separate(file, into) 
}

read_dir = function(path, pattern, into) {
  files = list.files(path = path,
                   pattern = pattern,
                   recursive = TRUE,
                   full.names = TRUE)
  plyr::ldply(files, my_read_csv, into = into)
}
```

This code can then be used with a call to the `read_dir` function which will 
read all the files in a particular directory (`path`) with a particular 
pattern (`pattern`) in the filename, split the directory and filename
structure into variables (`into`), and save the result as a `data.frame`.

For example, suppose there is a `bee/` data directory with the YYYY/MM/DD/Recorder
subdirectory structure and filenames HH_MM_site.csv. 
Then from the directory below `bee/`, running the following will read all the
`csv` files and return a `data.frame` with the variables in the data files as well as 
day, time, site, and recorder information.

```{r reading_files, eval=FALSE}
bee <- read_dir('bee', 'csv', into = c("bee", "year", "month", "day", "recorderID"
                                       "hour", "minute", "siteID", "csv")) %>%
  select(-bee, -csv)
```

The `read_dir` function is set up to extract all information including 
directory name, `bee/`, and file extension, `csv`, which will likely not be 
needed and thus the last line will remove these variables from the created 
`data.frame`.


## Joining scripts

As we are promoting a ``Store each piece of data only once'', 
we will likely need to combine multiple `data.frame`s. 
For example, suppose each site has a particular treatment and we are measure 
a response variable multiple times on each site. 
In this scenario, we will have (at least) two `data.frame`s:
one that contains the site-treatment information and another that contains 
the response variable at the sites at different times.
Suppose that these variables are called `site` and `response` respectively and
they have a common variable `siteID`. 
These two `data.frame`s can be combined using the following script.

```{r joining-script, eval=FALSE}
library("dplyr")
combined <- response %>% left_join(site, by="siteID")
```

There are additional joining functions in the `dplyr` package including `join`
`right_join`, `full_join`, `inner_join`, etc. 



## Summarizing scripts

## Analyzing scripts